|

### 2019-10-23 manuscript/manuscript.Rmd
decision tree plots

In this section, we attempt to develop a working definition of interpretability for ratemaking applications. While we will not provide a comprehensive survey of the prolific and fast evolving ML interpretability literature, we draw from it as appropriate in setting the stage for our discussion. There is not a standard definition of interpretability with consensus from all researchers; here are a few from frequently cited papers:

@@

We look at some examples of models that are considered by many to be interpretable. The GLM is probably the most oft-cited example of such a model (or rather, modeling framework) in the actuarial literature. Given a set of inputs, we can easily reason about what the output of the model is. As an illustrative example, we may have a claims severity model with age, amount of insurance (AOI), and gender as features; assuming a log link function, we have

\begin{equation}
  \log(E(\text{severity})) = \beta_0 + \beta_1 \cdot \text{age} + \beta_2 \cdot \text{AOI} + \beta_3 \cdot \text{gender}_{\text{male}}.
\end{equation}

Here, we can tell, for example, what the model would predict for the expected severity if we were to increase age by a certain amount, *all else being equal*, because the relationship between the feature and the response is simply the coefficient $\beta_1$ and the link function. Another commonly cited example of an interpretable model is a decision tree. An illustrative example is shown in (TODO: FIGURE). Here, the prediction is arrived at by following a sequence of if-else decisions.

@@

Similarly, one can see that in (TODO: FIGURE), large and deep trees are tough to reason about. In other words, even when working within the framework of an "interpretable" class of models, we may still end up with something that many would consider "black box."

@@

In Figure \@ref(fig:fi), we show a plot of variable importances. We see that FOO and BAR are the most important variables in terms of contributing accuracy to our model. Note that these measures do not provide information regarding the directional sensitivity of the predictors on the response. In our particular example, we see that the "make" variable contributes most to the accuracy of the model with "XXX" contributing the least. This provides a way for the audience to quickly glance at the most relevant variables, and ask further question as necessary. For example, if a variable that can potentially be a proxy for a discriminatory variable appears as very important, more investigation would be warranted.

>>>>

In this section, we attempt to develop a working definition of interpretability for ratemaking applications. While we will not provide a comprehensive survey of the prolific and fast evolving ML interpretability literature, we draw from it as appropriate in setting the stage for our discussion. Even among researchers in ML, there is not a consensus on the definition of interpretability; here are a few from frequently cited papers:

@@

Let us look at some examples of models that are considered by many to be interpretable. The GLM is probably the most oft-cited example of such a model (or rather, modeling framework) in the actuarial literature. Given a set of inputs, we can easily reason about what the output of the model is. As an illustrative example, we may have a claims severity model with driver age, amount of insurance (AOI), and sex as features; assuming a log link function, we have

\begin{equation}
  \log(E(\text{severity})) = \beta_0 + \beta_1 \cdot \text{age} + \beta_2 \cdot \text{AOI} + \beta_3 \cdot \text{sex}_{\text{male}}.
\end{equation}

Here, we can tell, for example, what the model would predict for the expected severity if we were to increase age by a certain amount, *all else being equal*, because the relationship between the feature and the response is simply the coefficient $\beta_1$ and the link function. Another commonly cited example of an interpretable model is a decision tree. An illustrative example is shown in Figure \@ref(fig:tree-plot1). Here, the prediction is arrived at by following a sequence of if-else decisions.

```{r tree-plot1, echo = FALSE, fig.cap="A simple decision tree for loss cost prediction.", out.width="70%", fig.align='center'}
knitr::include_graphics("figures/tree-plot1.png")
```

@@

Similarly, one can see that in Figure \@ref(fig:tree-plot2), large and deep trees are tough to reason about. In other words, even when working within the framework of an "interpretable" class of models, we may still end up with something that many would consider "black box."

```{r tree-plot2, echo = FALSE, fig.cap="A more complex decision tree for loss cost prediction.", out.width="90%", fig.align='center'}
knitr::include_graphics("figures/tree-plot2.png")
```

@@

In Figure \@ref(fig:fi), we show a plot of variable importances. We see that FOO and BAR are the most important variables in terms of contributing accuracy to our model. Note that these measures do not provide information regarding the directional sensitivity of the predictors on the response. In our particular example, we see that the "make" variable contributes most to the accuracy of the model with "sex" contributing the least. This provides a way for the audience to quickly glance at the most relevant variables, and ask further question as necessary. For example, if a variable that can potentially be a proxy for a discriminatory variable appears as very important, more investigation would be warranted.
